---
title: "Carbon Emissions Analysis of Commercial Office Buildings"
subtitle: "Prepared for the Kansas City Office of Sustainability"
author: "Colin Haley"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# install.packages("readxl")  # uncomment if not already installed
library(readxl)
```

# Introduction and Background

The Kansas City Office of Sustainability is evaluating policies to reduce carbon dioxide emissions from large commercial office buildings. The city is specifically interested in understanding which building characteristics are most closely associated with annual emissions, and whether energy retrofits change the way building size affects emissions. This report summarizes an analysis of 120 office buildings, using their carbon emissions data to answer these questions and to inform policy recommendations.

**Research Questions**

1. Which characteristics of commercial buildings are most closely associated with carbon emissions?
2. Is an energy retrofit an effective way to reduce emissions when compared with other building characteristics?

# Data Description

The dataset contains 120 office buildings and includes one response variable and five predictor variables.

**Response Variable**

- `Emissions_tons`: annual carbon dioxide emissions, in metric tons. Quantitative, continuous.

**Predictor Variables**

- `FloorArea_kSqFt`: building size, in thousands of square feet. Quantitative, continuous.
- `EnergyStarScore`: energy efficiency rating. Quantitative, continuous.
- `RetrofitStatus`: whether the building has undergone an energy retrofit. Categorical, nominal. A dummy variable (`RetrofitStatusRetrofit`) is created for this predictor, with "Retrofit" used as the reference category.
- `TransitScore`: public transit accessibility score. Quantitative, continuous.
- `OccupancyRate`: average annual occupancy rate. Quantitative, continuous.

```{r load-data}
# Update this path to point to your local copy of the data file
building_emissions_sustainability_Final <- read_excel("building_emissions_sustainability_Final.xlsx")
```

# Data Preprocessing

Before analysis, the data were cleaned and prepared as follows:

- Missing and null values were removed.
- Variables were converted to appropriate data types.
- The categorical predictor, `RetrofitStatus`, was converted into a dummy variable for use in the regression models.

```{r preprocess}
building_emissions_sustainability_Final <- na.omit(building_emissions_sustainability_Final)
building_emissions_sustainability_Final$RetrofitStatus <- as.character(building_emissions_sustainability_Final$RetrofitStatus)
```

# Exploratory Data Analysis

```{r summary-stats}
summary(building_emissions_sustainability_Final)
```

```{r std-devs}
sd(building_emissions_sustainability_Final$Emissions_tons)
sd(building_emissions_sustainability_Final$FloorArea_kSqFt)
sd(building_emissions_sustainability_Final$EnergyStarScore)
sd(building_emissions_sustainability_Final$TransitScore)
sd(building_emissions_sustainability_Final$OccupancyRate)
```

# Scatterplots and Visual Analysis

Scatterplots were generated to visually assess the relationship between emissions and each numerical predictor.

## Emissions by Square Footage: Strong Positive Relationship

```{r scatter-floorarea}
plot(building_emissions_sustainability_Final$Emissions_tons,
     building_emissions_sustainability_Final$FloorArea_kSqFt,
     main = "Emissions by Square Footage", xlab = "Emissions",
     ylab = "Square Footage", col = "blue")
```

## Emissions by Energy Star Score: Weak Negative Relationship

```{r scatter-energystar}
plot(building_emissions_sustainability_Final$Emissions_tons,
     building_emissions_sustainability_Final$EnergyStarScore,
     main = "Emissions by Energy Star Score", xlab = "Emissions",
     ylab = "Energy Star Score", col = "blue")
```

## Emissions by Transit Score: Weak Positive Relationship

```{r scatter-transit}
plot(building_emissions_sustainability_Final$Emissions_tons,
     building_emissions_sustainability_Final$TransitScore,
     main = "Emissions by Transit Score", xlab = "Emissions",
     ylab = "Transit Score", col = "blue")
```

## Emissions by Occupancy Rate: Weak Positive Relationship

```{r scatter-occupancy}
plot(building_emissions_sustainability_Final$Emissions_tons,
     building_emissions_sustainability_Final$OccupancyRate,
     main = "Emissions by Occupancy Rate", xlab = "Emissions",
     ylab = "Occupancy Rate", col = "blue")
```

# Correlation Analysis

Pairwise correlations were calculated between emissions and each numerical predictor to quantify the strength and direction of each relationship.

```{r correlations}
# Correlation between emissions and square footage
cor(building_emissions_sustainability_Final$Emissions_tons,
    building_emissions_sustainability_Final$FloorArea_kSqFt)

# Correlation between emissions and Energy Star Score
cor(building_emissions_sustainability_Final$Emissions_tons,
    building_emissions_sustainability_Final$EnergyStarScore)

# Correlation between emissions and transit score
cor(building_emissions_sustainability_Final$Emissions_tons,
    building_emissions_sustainability_Final$TransitScore)

# Correlation between emissions and occupancy rate
cor(building_emissions_sustainability_Final$Emissions_tons,
    building_emissions_sustainability_Final$OccupancyRate)
```

# Simple Linear Regression

A simple linear regression model was fit using annual emissions as the response variable and building size as the sole predictor.

**Regression equation:** Yhat = 21.9775 + 0.60963 × (building size)

```{r slr}
slr <- lm(Emissions_tons ~ FloorArea_kSqFt,
          data = building_emissions_sustainability_Final)
summary(slr)
```

**Slope interpretation:** for every one square foot increase in building floor area, total emissions are estimated to increase by 0.60963 tons, holding all else constant.

**R-squared interpretation:** the R-squared value for this model is 0.364, meaning that 36.4% of the variation in emissions can be explained solely by the square footage of the building being analyzed.

**Model sufficiency for policy decisions:** this simple linear regression model is not sufficient on its own for sustainability policy decisions. Because the model relies on a single predictor, it explains only 36.4% of the variability in emissions. Additional predictors are needed to build a more reliable model that captures more of the variation in emissions.

# Multiple Regression: Main Effects Model

A multiple regression model was fit using annual emissions as the response variable and all available predictors, referred to as the main effects model.

**Regression equation:** Yhat = 70.937 + 0.70045 × (square footage) − 0.61628 × (Energy Star Score) − 69.88 × (No Retrofit) − 0.18029 × (transit score) + 0.28365 × (occupancy rate)

```{r main-effects-model}
RetrofitDummy <- model.matrix(~RetrofitStatus,
                               data = building_emissions_sustainability_Final)

Model1 <- lm(Emissions_tons ~ FloorArea_kSqFt + EnergyStarScore +
               RetrofitStatus + TransitScore + OccupancyRate,
             data = building_emissions_sustainability_Final)
summary(Model1)
```

**Coefficient interpretations**

- Square footage: for each one unit increase in square footage, total emissions increase by 0.70045 tons, holding other variables constant.
- Energy Star Score: for each one unit increase in Energy Star Score, total emissions decrease by 0.61628 tons, holding other variables constant.
- Retrofit status: a retrofitted building is associated with a 69.88 ton decrease in total emissions compared with a building that has not been retrofitted.
- Transit score: for each one unit increase in transit score, total emissions decrease by 0.18029 tons, holding other variables constant.
- Occupancy rate: for each one unit increase in occupancy rate, total emissions increase by 0.28365 tons, holding other variables constant.
- Intercept: if all other predictors were zero, total emissions would equal 70.93687 tons. This value has limited practical use, since it does not correspond to any real building.

**Statistical significance:** square footage, Energy Star Score, and retrofit status are all statistically significant predictors at the 0.05 significance level.

**Effect of retrofit status:** after controlling for the other variables in the model, retrofit status continues to be associated with a meaningful reduction in emissions. Using the dummy variable created for this predictor, with Retrofit as the reference category, the model estimates that retrofitted buildings produce approximately 69.88 fewer tons of emissions than buildings that have not been retrofitted.

# Interaction Model: Does Retrofit Moderate the Size Effect?

The city is especially interested in whether the effect of building size on emissions differs for retrofitted buildings. To evaluate this, a multiple regression model was fit that includes all available predictors, retrofit status, and an interaction term between floor area and retrofit status.

**Regression equation:** Yhat = 37.30354 + 1.12474 × (square footage) + 36.99 × (retrofit) − 0.62373 × (Energy Star Score) − 0.03592 × (transit score) − 0.05466 × (occupancy rate) − 0.79772 × (interaction between square footage and retrofit)

```{r interaction-model}
Model2 <- lm(Emissions_tons ~ FloorArea_kSqFt * RetrofitStatus +
               EnergyStarScore + TransitScore + OccupancyRate,
             data = building_emissions_sustainability_Final)
summary(Model2)
```

**Interpretation of key coefficients**

- Square footage: for each one unit increase in square footage, total emissions increase by 1.12474 tons, holding other variables constant.
- Retrofit status: for a retrofitted building, emissions increase by 36.99425 tons compared with a building that has not been retrofitted, before accounting for the interaction with square footage.
- Interaction between retrofit and square footage: for retrofitted buildings, the slope of the square footage relationship is reduced by 0.79772 compared with non-retrofitted buildings, meaning that the emissions benefit of a retrofit grows as building size increases.

**Statistical significance of the interaction:** the interaction term has a p-value of 1.71e-12, which is statistically significant at the 0.05 significance level.

**Implications for retrofit value by building size:** the negative coefficient of −0.79772 on the interaction term indicates that a retrofit has an increasingly positive effect on reducing emissions as square footage increases. In other words, the larger the building, the more emissions a retrofit is likely to reduce compared with a smaller building.

# Policy Interpretation of Retrofits

Using the interaction model, the marginal effect of a retrofit was calculated for a building with a floor area of 150,000 square feet (`FloorArea_kSqFt = 150`).

```{r marginal-effect}
marginal_retrofit_effect <- 36.99425 + (-0.79772 * 150)
marginal_retrofit_effect
```

**Interpretation:** based on this calculation, applying a retrofit to a 150,000 square foot building would be expected to decrease emissions by approximately 82.66 tons.

**Influence on city retrofit incentives:** a reduction of more than 80 tons of emissions for a large building of this size should strongly influence the city to incentivize retrofit programs. This finding shows that retrofits deliver substantially greater emissions reductions in large buildings, which makes incentive programs targeted at large buildings a particularly effective use of city resources.

# Prediction and Uncertainty

Using the interaction model, annual carbon dioxide emissions were predicted for a building with the following characteristics: floor area of 150,000 square feet, Energy Star Score of 80, retrofit status of "Yes," transit score of 85, and occupancy rate of 90 percent.

```{r prediction-manual}
predicted_emissions <- 37.30354 + (1.12474 * 150) + (36.99425 * 1) +
  (-0.62373 * 80) + (-0.03592 * 85) + (-0.05466 * 90) + (-0.79772 * 150 * 1)
predicted_emissions
```

This can also be confirmed using `predict()` on the fitted interaction model:

```{r prediction-predict}
new_building <- data.frame(
  FloorArea_kSqFt = 150,
  RetrofitStatus = "Retrofit",
  EnergyStarScore = 80,
  TransitScore = 85,
  OccupancyRate = 90
)
predict(Model2, newdata = new_building)
```

The predicted value is approximately 65.48 tons of annual emissions.

# Conclusion and Recommendations

Across the models tested, square footage and retrofit status emerged as the two strongest drivers of carbon emissions among commercial office buildings in this dataset. Square footage was consistently associated with higher emissions, while retrofit status was consistently associated with lower emissions, and the interaction model shows that this emissions reduction becomes larger as building size increases.

**Retrofit effectiveness:** retrofits have a positive impact on reducing carbon emissions, and this impact scales with building size. Larger buildings see a greater emissions benefit from retrofit upgrades than smaller buildings, where retrofits appear to be less effective in reducing emissions.

**City planning recommendations**

1. For buildings larger than 100,000 square feet, prioritize retrofit programs, since retrofits produce the greatest emissions reductions in this size range.
2. For smaller buildings, focus policy efforts on reducing square footage or improving space efficiency rather than mandating retrofits, since the emissions benefit of retrofits is smaller at this scale.
3. Consider using Energy Star Score as an additional variable when identifying buildings most likely to reduce emissions, since higher scores are associated with lower emissions.
